---
title: "CBM_core Manual"
subtitle: "v.`r SpaDES.core::moduleMetadata(module = 'CBM_core', path = '..')$version`"
date: "Last updated: `r Sys.Date()`"
output:
  md_document:
    pandoc_args: ["--wrap=none"]
    variant: markdown-raw_html
editor_options: 
  chunk_output_type: console
---
```{r setup-CBM_core, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE, 
  message = FALSE, 
  warning = FALSE
)
```
```{r format-metadata, include=FALSE}
md_table <- function(mdTable){

  names(mdTable)[names(mdTable) == "objectName"]  <- "Name"
  names(mdTable)[names(mdTable) == "objectClass"] <- "Class"
  names(mdTable)[names(mdTable) == "desc"]        <- "Description"

  if ("columns" %in% names(mdTable)){
    mdTable$`Table columns` <- sapply(mdTable$columns, function(columns){
      if (length(columns) > 0){
        paste(paste0("*", names(columns), "*: ", columns), collapse = "<br>")
      }else ""
    })
  }

  mdTable[, intersect(c("Name", "Class", "Description", "Table columns"), names(mdTable))]
}
```

# CBM_core

:::{.rmdimportant}
This documentation is a work in progress. Potential discrepancies and omissions may exist for the time being. If you find any, please contact us [here]("https://github.com/PredictiveEcology/CBM_core/issues").
:::

## Overview

[CBM_core](https://github.com/PredictiveEcology/CBM_core) is the central module of [spadesCBM](https://github.com/PredictiveEcology/spadesCBM) implementing the [Carbon Budget Model of the Canadian Forest Sector](https://natural-resources.canada.ca/climate-change/climate-change-impacts-forests/carbon-budget-model) version 4 (CBM-CFS4). This is where carbon transfers are calculated at every time step, disturbances are applied, and stocks are tracked.

The [CBM4r](https://github.com/PredictiveEcology/CBM4r) R package provides an R interface to CBM4 Python applications.
The [CBMutils](https://github.com/PredictiveEcology/CBMutils) R package contains tools to plot simulation results.

The SpaDES toolkit enables a modular, repeatable, and continuous workflow. This brings the transparency and flexibility needed for scientists to modify, evaluate, and test new inputs, new data, and new algorithms while also providing accessibility to non-researcher users.

## Module inputs

The simulation study area is set as a raster grid (input `masterRaster`). 
In *CBM_core*, each forested `masterRaster` pixel represents a "stand" (input `standDT`) with one or more cohorts.
Cohorts (input `cohortDT`) share the same age (time since disturbance) and other classifiers (such as tree species) associating them to growth curves (inputs `gcMeta` and `gcIncrements`). Disturbance events (inputs `disturanceMeta` and `disturbanceEvents`) occur within stands throughout the simulation affecting their cohorts.

CBM parameters can be altered by providing a customized version of the CBM defaults SQLite database (input `cbm_defaults_db`).

```{r moduleInputs-CBM-core, results = 'asis', comment = ''}
SpaDES.core::moduleInputs("CBM_core", "..") |>
  md_table() |>
  knitr::kable(format = "pipe", caption = "*CBM_core* input objects") |> 
  print()
```

```{r moduleParams-CBM-core, results = 'asis', comment = ''}
mdTable <- SpaDES.core::moduleParams("CBM_core", "..")
mdTable <- mdTable[, c("paramName", "paramClass", "default", "paramDesc")]
names(mdTable) <- c("Name", "Class", "Default", "Description")

mdTable |>
  knitr::kable(format = "pipe", caption = "*CBM_core* input parameters") |> 
  print()
```

## Module events

Each of the following events are run for the first simulation year. The `annualDisturbances` and `annualStep` events are repeated for each following simulation year. 

### `Init`

This event sets the CBM4 data directory path (output `CBM4data`) and links *CBM4r* to the CBM parameters database (input `cbm_defaults_db`) if provided.

The `r-CBM4` Python virtual environment is set up and activated. If a suitable version of Python is not available, it will be installed.

### `setStands`

This event reads the input stand table (input `standDT`) and adds additional attributes required throughout the simulation. See [`CBM4r::cbm4_set_grid_meta`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_set_grid_meta-cbm4-set-grid-metadata) for more details.

### `spinup`

This event initializes the landscape by performing the CBM spinup.

The cohort inventory (input `cohortDT`) is written the CBM4 data directory with [`CBM4r::cbm4_write_inventory`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_write_inventory-cbm4-write-inventory). 
Growth increments (inputs `gcMeta` and `gcIncrements`) and ecological parameters (input `cbm_defaults_db`) are written to the CBM4 data directory with [`CBM4r::cbm4_write_spinup_parameters`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_write_spinup_parameters-cbm4-write-spinup-parameters).
The spinup is performed with [`CBM4r::cbm4_spinup`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_spinup-cbm4-spinup).

If parameter `fixedCohorts = FALSE` the cohort inventory and pools state will be read into the `cohortDT` object for use by other modules.

### `annualDisturbances`

This event writes disturbance events for the current simulation year to the CBM4 data directory to be used in the `annualStep`. This process is repeated yearly to allow for disturbances to be provided by other modules on a yearly basis.

Disturbances are written to file with [`CBM4r::cbm4_write_disturbance`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_write_disturbance-cbm4-write-disturbance).

### `annualStep`

This event runs the CBM annual step where carbon transfers are applied for each simulation year.

The cohort inventory and pools state will be read directly from the CBM4 data directory unless parameter `fixedCohorts = FALSE` in which case the cohort inventory and pools state (input `cohortDT`) will be written the CBM4 data directory with [`CBM4r::cbm4_write_simulation_inventory`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_write_simulation_inventory-cbm4-write-simulation-inventory).

Growth increments (inputs `gcMeta` and `gcIncrements`) and ecological parameters (input `cbm_defaults_db`) are written to the CBM4 data directory with [`CBM4r::cbm4_write_step_parameters`](https://github.com/PredictiveEcology/CBM4r/blob/main/inst/docs/CBM4r_Reference_Manual.md#cbm4_write_step_parameters-cbm4-write-step-parameters).

If parameter `fixedCohorts = FALSE` the updated cohort inventory and pools state will be read into the `cohortDT` object for use by other modules.

### `plot`

This final optional event is where all plotting occurs. 
Plots will be saved as `.png` files in the project outputs folder.


## Module outputs

The primary output produced by *CBM_core* is the CBM4 data directory containing simulation data in the [CBM4 spatial dataset](https://github.com/cat-cfs/tech_docs/tree/main/dev/cbm4/structure#cbm4-spatial-datasets) format.
Simulation results can either be read directly as Parquet tables or more easily with the [CBM4 results processor](https://github.com/cat-cfs/tech_docs/tree/main/dev/cbm4/structure#results-processing).
For more information see the [CBM4r documentation](https://github.com/PredictiveEcology/CBM4r).

```{r moduleOutputs-CBM-core, results = 'asis', comment = ''}
mdTable <- SpaDES.core::moduleOutputs("CBM_core", "..") |>
  md_table() |>
  knitr::kable(format = "pipe", caption = "*CBM_core* output objects") |> 
  print()
```

### Plotting outputs with CBMutils

```{r plotOutputs-CBM-core, echo = TRUE, eval = FALSE}

# Read results as CBM4 SQLResultsProcessor
cbm4_results <- CBM4r::cbm4_results_processor(sim$CBM4data)

# Plot emissions and products
CBMutils::cbm4PlotEmissionsProducts(cbm4_results, yearStart = start(sim))

# Plot pool proportions
CBMutils::cbm4PlotPoolProportions(cbm4_results, yearStart = start(sim))

# Plot Net Primary Productivity (NPP)
plotNPP <- CBMutils::cbm4MapNPP(cbm4_results, yearStart = start(sim))
plotNPP[[as.character(start(sim))]]
plotNPP[[as.character(end(sim))]]

# Plot total carbon
plotTC <- CBMutils::cbm4MapTotalCarbon(cbm4_results, yearStart = start(sim))
plotTC[[as.character(start(sim))]]
plotTC[[as.character(end(sim))]]

```

## Links to other modules

- [CBM_defaults](https://github.com/PredictiveEcology/CBM_defaults)
- [CBM_dataPrep](https://github.com/PredictiveEcology/CBM_dataPrep)
- [CBM_dataPrep_SK](https://github.com/PredictiveEcology/CBM_dataPrep_SK)
- [CBM_vol2biomass](https://github.com/PredictiveEcology/CBM_vol2biomass)


